---
title: "C03_assignment"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("setup.R")

SPECIES2 <- "Malacoraja senta"
VERSION2 <- "v1"
```

## Species 2 Malacoraja senta

This is similar to the thorny skate but is only really present in the Gulf of Maine
so it could be cool to look at. I then loaded in the brickman data, created a tibble of the monthly counts and plotted them on the brickman mask.


```{r load_obs_mask}
coast = read_coastline()
obs = read_observations(scientificname = SPECIES2, minimum_year = 1970)
db = brickman_database() |>
  filter(scenario == "STATIC", var == "mask")
mask = read_brickman(db)
```

```{r all_observations_count}
LON0 = -67
LAT0 = 46
all_counts = count(st_drop_geometry(obs), month) # counting is faster without spatial baggage
all_counts
```

```{r all_observations}
ggplot() +
  geom_sf(data = obs, alpha = 0.2, shape = "circle small", size = 1) +
  geom_sf(data = coast, col = "orange") +
  geom_text(data = all_counts,
            mapping = aes(x = LON0, 
                          y = LAT0, 
                          label = sprintf("n: %i", .data$n)),
                          size = 3) + 
  labs(x = "Longitude", y = "Latitude", title = "All observations") +
  facet_wrap(~month)
```

## Thinning observations (one obs per grid)
```{r thin_observations}
thinned_obs = sapply(month.abb,
               function(mon){ 
                 thin_by_cell(obs |> filter(month == mon), mask)
               }, simplify = FALSE) |>
  dplyr::bind_rows() 

# another count
thinned_counts = count(st_drop_geometry(thinned_obs), month)

ggplot() +
  geom_sf(data = thinned_obs, 
          alpha = 0.2, 
          shape = "circle small", 
          size = 1) +
  geom_sf(data = coast, col = "orange") +
  geom_text(data = thinned_counts,
            mapping = aes(x = LON0, 
                          y = LAT0, 
                          label = sprintf("n: %i", .data$n)),
                          size = 3) + 
  labs(x = "Longitude", y = "Latitude", title = "Thinned observations") +
  facet_wrap(~month)
```

## Bias map
```{r bias_map}
bias_map = rasterize_point_density(obs, mask) # <-- note the original observations

ggplot() +
  geom_stars(data = bias_map, aes(fill = count)) +
  scale_fill_viridis_b(na.value = "transparent") +
  geom_sf(data = coast, col = "orange") + 
  labs(x = "Longitude", y = "Latitude", title = "Bias map using all observations")
```


## Creating background points
```{r n_background_per_month}
nback_avg = 1000
nback_avg
```

```{r sample_background}
obsbkg = sapply(month.abb,
    function(mon){ 
      sample_background(thinned_obs |> filter(month == mon), # <- just this month
                       bias_map,
                       method = "bias",  # <-- it needs to know it's a bias map
                       return_pres = TRUE, # <-- give me the obs back, too
                       n = nback_avg) |>   # <-- how many points
        mutate(month = mon, .before = 1)
    }, simplify = FALSE) |>
  bind_rows() |>
  mutate(month = factor(month, levels = month.abb))
obsbkg 
```

```{r nback_avg}
# drop the spatial baggage to speed the tally
count(st_drop_geometry(obsbkg), month, class)
```
## Background vs Presense Plot
```{r plot_presence_background}
ggplot() +
  geom_sf(data = dplyr::filter(obsbkg, class == "background"),
          aes(color = class),
          alpha = 0.2, size = 1) +
  geom_sf(data = dplyr::filter(obsbkg, class == "presence"),
          aes(color = class),
          alpha = 0.9, size = 1) +
  geom_sf(data = coast, col = "orange") +
  scale_color_okabe_ito() +
  labs(x = "Longitude", y = "Latitude", title = "Presence vs Background") +
  theme_bw() +
  facet_wrap(~month)
```

## Saving model input file
```{r save_model_input}
write_model_input(obsbkg, scientificname = SPECIES2, version = VERSION2)
```

```{r read_model_input}
x = read_model_input(scientificname = SPECIES2, version = VERSION2)
```

```{r read_brickman}
db = brickman_database() |>
  dplyr::filter(scenario == "PRESENT", interval == "mon")
present = read_brickman(db)
```

## Pair plot of present
```{r pairs}
pairs(present)
```

```{r filter_collinear}
keep = filter_collinear(present, method = "cor_caret", cutoff = 0.65)
keep
```

```{r drop_me}
drop_me = attr(keep, "to_remove")
drop_me
```

```{r add_depth_and_month}
keep = c("depth", "month", "SST", "Tbtm", "MLD")
keep
```

```{r present_model_input}
model_input = read_model_input(scientificname = SPECIES2, version = VERSION2)

model_input
```
## Violin plot of kept variables

```{r reload_brickman}
present = read_brickman(add = c("depth"))

variables = extract_brickman(present, model_input, form = "wide")
variables

variables = variables |>
  mutate(class = model_input$class) |>
  select(-.id)

variables = variables |>
  select(all_of(c("geom", "class", keep)))

variables

plot_pres_vs_bg(variables |> select(-month), "class")
```


```{r make_config}
cfg = list(
  version = VERSION2,
  scientificname = SPECIES2,
  background = "average of observations per month",
  keep_vars =  keep)
```

```{r save config}
ok = make_path(data_path("models")) # make a directory for models
write_configuration(cfg)            
```
```{r save_input_data}
write_model_input(variables, scientificname = SPECIES2, version = VERSION2)
```





