---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Species 2 Malacoraja senta Model Building (assignment 4)

This is similar to the thorny skate but is only really present in the Gulf of Maine so it could be cool to look at. I then loaded in the brickman data, created a tibble of the monthly counts and plotted them on the brickman mask.

```{r}
source("setup.R")

library(workflowsets)
library(tune)


SPECIES2 <- "Malacoraja senta"
VERSION2 <- "v1"
```

## Loading data
```{r load_data}
cfg = read_configuration(scientificname = SPECIES2, version = VERSION2)

model_input = read_model_input(
  scientificname = SPECIES2,
  version = VERSION2,
  log_me = c("depth")
) |>
  dplyr::mutate(
    month  = month_as_number(.data$month),
    weight = dplyr::if_else(class == "presence", 5, 1)
  ) |>
  dplyr::select(all_of(c("class", "weight", cfg$keep_vars)))

```

## splitting into training and testing data
```{r initial_split}
model_input_split = spatial_initial_split(model_input, 
                        prop = 1 / 5,     # 20% for testing
                        strategy = spatial_block_cv) # see ?spatial_block_cv
model_input_split
```

```{r initial_split_plot}
autoplot(model_input_split)
```

```{r cv_training}
tr_data = training(model_input_split)
cv_tr_data <- spatial_block_cv(tr_data,
  v = 5,     
  cellsize = grid_cellsize(model_input),
  offset = grid_offset(model_input) + 0.00001
)
autoplot(cv_tr_data)
```

## creating the recipe
```{r recipe}
# recipe should be defined from the FULL training data so roles are stable
rec = recipe(class ~ ., data = tr_data) |>
  # make sure weight is not a predictor
  update_role(weight, new_role = "case_weight") |>
  # make sure geom is not a predictor
  update_role(geom, new_role = "coords")

rec
summary(rec)
```

```{r make_workflow}
wflow = workflow_set(
  
  preproc = list(default = rec), # not much happening in our preprocessor
  
  models = list(                 # but we have 4 models to add
    
      # very simple - nothing to tune
      glm = logistic_reg(
          mode = "classification") |>
        set_engine("glm"),
      
      # two knobs to tune
      rf = rand_forest(
         mtry = tune(),
         trees = 1000,
         mode = "classification"
      ) |>
         set_engine(
            "ranger",
            importance = "impurity"
      ),

      
      # so many things to tune!
      btree = boost_tree(
          mtry = tune(), 
          trees = tune(), 
          tree_depth = tune(), 
          learn_rate = tune(), 
          loss_reduction = tune(), 
          stop_iter = tune(),
          mode = "classification") |>
          set_engine("xgboost"),
    
      # just two again
      maxent = maxent(
          feature_classes = tune(),
          regularization_multiplier = tune(),
          mode = "classification") |>
        set_engine("maxnet")
  )
)
wflow
```

```{r metrics}
metrics = sdm_metric_set(yardstick::roc_auc)
metrics
```

```{r fit, warning = FALSE}
wflow <- wflow |>
  workflow_map("tune_grid",
    resamples = cv_tr_data, 
    grid = 5,
    metrics = metrics, 
    verbose = TRUE)
```

## Plots of model comparisons
Here's a quick plot of the outputs, with each metric being showcased.
```{r plot_wflow}
#autoplot didnt work so i just ggplot plotted it instead
metrics_tbl <- tune::collect_metrics(wflow)

metrics_tbl |>
  dplyr::filter(.metric == "roc_auc") |>
  ggplot2::ggplot(ggplot2::aes(x = mean, y = wflow_id)) +
  ggplot2::geom_point(size = 3) +
  ggplot2::labs(
    x = "Mean ROC AUC",
    y = "Model",
    title = "Model performance comparison"
  )
```


```{r select_best}
model_fits = workflowset_selectomatic(wflow, model_input_split,
                                  filename = "Malacoraja_senta-v1-model_fits",
                                  path = data_path("models"))
model_fits
```

```{r model_fit_metrics}
model_fit_metrics(model_fits)
```


```{r model_fit_confmat}
model_fit_confmat(model_fits)
```

```{r model_fit_roc_auc}
model_fit_roc_auc(model_fits)
```

```{r model_fit_vip}
model_fit_varimp_plot(model_fits)
```

```{r boosted tree}
rf = model_fits |>
  filter(wflow_id == "default_btree")
rf
```

```{r rf_splits}
autoplot(rf$splits[[1]])
```

```{r rf_metrics}
rf$.metrics[[1]]
```

```{r rf_preds}
rf$.predictions[[1]]
```

```{r rf_workflow}
rf$.workflow[[1]]
```

```{r pd_plot}
# pretty sure this doesnt work because there is a 5-6 variable mismatch
# model_fit_pdp(model_fits, wid = "default_btree", title = "Boosted Tree")
```





## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

